\OurTool{}: Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles

### 2019-06-21 IPANEMAP-2019.tex
Updates from Overleaf

\paragraph{Stops-inducing probing protocol (SHAPE-CE).} 
6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature during 10 minutes in the probing buffer (40 mM HEPES at 7.5 pH, 100 mM KCl, in presence or absence of 5 mM MgCl$_\text{2}$. After a 10 minutes incubation at 37\degree{}C, RNAs were treated with 2 mM of SHAPE reagent or DMSO (negative control) and incubated for 2 (BzCN), 5 (1M7), 30 (NMIA) or 60 (NAI) minutes at 37\degree{}C. Modified or unmodified RNAs were purified by ethanol precipitation and pellets were resuspended in 10 $\mu$l of water.
Modifications were revealed by reverse transcription using 5’ fluorescently labelled primers (D2 or D4 WellRED, Sigma Aldrich 5’ CTG-TGA-ACT-AAT-GCT-GTC-CTT-TAA 3’) and M-MLV RNAse H- reverse transcriptase (Promega) as previously described~\cite{Deforges2017}, with only minor modification of the originally described SHAPE protocol~\cite{Wilkinson2006}. cDNAs were separated by capillary electrophoresis (Beckman Coulter, Ceq8000). Data were analyzed using the  QuSHAPE~\cite{Karabiber2013} software. RNA probing was performed in triplicate with distinct RNA preparations.

@@

For 1M7ILU3, the denaturation step was repeated three times : After a 1 minute incubation at 95\degree{}C 10 mM of 1M7 was added and the mix was incubated 1 minute at 95\degree{}C. This step was repeated 3 times. For 1M7, 1M7Mg, NMIA and NMIAMg, experiments were conducted as described above, except that the NGS library preparation was adapted to Ion Torrent sequencing. Sequences were mapped and analyzed with \Software{ShapeMapper2} for 1M7ILU and 1M7ILU3. For the other conditions they are mapped with a script based on \Software{ShapeMapper}, but adapted to Ion Torrent output files.

@@

Compared to bi-probing, MCC values achieved by integrating triplets of conditions now indicate a clear gain over the average performances of individual conditions, supporting a notion of complementarity between experiments. Indeed tri-probing experiments lead to an average +1.7\% MCC improvement (median +2.6) over the average of mono-probing prediction in the triplet. The gain of quality is widespread, and observed for 65\% of the triplets. Finally, 39 triplets were found to induce MCC values greater than 83\%.

@@

We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng dataset~\cite{Cheng2017}, which we supplemented with the original wild-type sequence (WT).
For each set of probing profiles, we executed \OurTool{} with default parameters, using a sample size of 1~000 structures. 
We also analyzed the WT sequence alone, reproducing the analysis 100 times to compare the variability across sets of mutants to the one induced by stochastic aspects of our method. 
Finally, we performed a restricted analysis focusing on the WT, supplemented with 2 probing profiles selected from the 20 most similar conditions to the WT , as assessed by the ensemble distance metrics. By comparing our results with those obtained for unrestricted pairs of mutants, we assess the influence of structure variability on the quality of our predictions. We report in Figure~\ref{fig:variantanalysis} the distribution of MCC values associated with the first prediction returned by \OurTool.

>>>>

\paragraph{Stops-inducing probing protocol (SHAPE-CE).} 
6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature during 10 minutes in the probing buffer (40 mM HEPES at 7.5 pH, 100 mM KCl, in presence or absence of 5 mM MgCl$_\text{2}$. After a 10 minutes incubation at 37\degree{}C, RNAs were treated with 2 mM of SHAPE reagent or DMSO (negative control) and incubated for 2 (BzCN), 5 (1M7), 30 (NMIA) or 60 (NAI) minutes at 37\degree{}C. Modified or unmodified RNAs were purified by ethanol precipitation and pellets were resuspended in 10 $\mu$l of water.
Modifications were revealed by reverse transcription using 5’ fluorescently labelled primers (D2 or D4 WellRED, Sigma Aldrich 5’ CTG-TGA-ACT-AAT-GCT-GTC-CTT-TAA 3’) and M-MLV RNAse H- reverse transcriptase (Promega) as previously described~\cite{Deforges2017}, with only minor modification of the originally described SHAPE protocol~\cite{Wilkinson2006}. cDNAs were separated by capillary electrophoresis (Beckman Coulter, Ceq8000). Data were analyzed using the  \Software{QuSHAPE}~\cite{Karabiber2013} software. RNA probing was performed in triplicate with distinct RNA preparations. The corresponding datasets are named after the probe followed by "-CE"

@@

For 1M7ILU3, the denaturation step was repeated three times : After a 1 minute incubation at 95\degree{}C 10 mM of 1M7 was added and the mix was incubated 1 minute at 95\degree{}C. This step was repeated 3 times(3D). For 1M7, 1M7Mg, NMIA and NMIAMg, experiments were conducted as described above, except that the NGS library preparation was adapted to Ion Torrent sequencing. Sequences were mapped and analyzed with \Software{ShapeMapper2} for 1M7ILU and 1M7ILU3. For the other conditions they are mapped with a script based on \Software{ShapeMapper}, but adapted to Ion Torrent output files.

@@

Compared to bi-probing, MCC values achieved by integrating triplets of conditions now indicate a clear gain over the average performances of individual conditions, supporting a notion of complementary between experiments. Indeed tri-probing experiments lead to an average +1.7\% MCC improvement (median +2.6) over the average of mono-probing prediction in the triplet. The gain of quality is widespread, and observed for 65\% of the triplets. Finally, 39 triplets were found to induce MCC values greater than 83\%.

@@

We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng dataset~\cite{Cheng2017}, which we supplemented with the original wild-type sequence (WT).
For each set of probing profiles \bsi{c'est ce qui correspondrait chez nous à du DMS Map mg ilu ce qu'il a fait. Tu peux l'insérer la avec notre nomenclature entre paranthèse step ?}, we executed \OurTool{} with default parameters, using a sample size of 1~000 structures. 
We also analyzed the WT sequence alone, reproducing the analysis 100 times to compare the variability across sets of mutants to the one induced by stochastic aspects of our method. 
Finally, we performed a restricted analysis focusing on the WT, supplemented with 2 probing profiles selected from the 20 most similar conditions to the WT , as assessed by the ensemble distance metrics. By comparing our results with those obtained for unrestricted pairs of mutants, we assess the influence of structure variability on the quality of our predictions. We report in Figure~\ref{fig:variantanalysis} the distribution of MCC values associated with the first prediction returned by \OurTool.
